### 2025-01-28 articleSummaries.qmd
Updated article summaries

---
title: "Article Summaries"
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
bibliography: references.bib
---

[@ziadi2023]This article focused on the length of productive life (LPL in the article) for female Floridian goats. The article discusses the failures of linear models and random regression linear models when it comes to modeling the LPL. As such, Cox proportional hazards model was chosen because it can handle censored records, time-dependent variables, and skewed distribution of characteristics. The does were classified as censored if they were living and uncensored if they were deceased. A preliminary analysis using likelihood ratio tests on a Weibull hazard survival model. This showed that the effects selected all had a significant effect on LPL. The Cox model is then provided for this specific problem. The results were that the age of the first kidding did significant influence LPL. The article provides tables of their results.

[@zapletal2021] The goal of this paper was to examine how explanatory variables influence on an event and if the estimates of the effects of covariates will differ based on the approach that was selected. This paper in particular focused on the Cox proportional hazard and the competing risks approach. The Cox proportional hazards model provides an unspecified baseline hazard function which is beneficial when dealing with a single outcome of interest or one that does not have a specific cause (equations are provided in article). The article describes which competing risks model is the one that will be used in the modeling section. The predictors used by the author were gender, age when the contract began, and region where the person lived. The particular data used in this article was from a Czech Republic insurance company who provide policies for certain illness events until the event occurs. The insured will receive money if the event occurs, but their contract is then complete. Cox models have an assumption of proportional hazards which means the ratios of the hazards need to be independent of time to use the model. The test used was the zph test using Schoenfield residuals (paper is linked in article) and the hypothesis was rejected for age, but not for gender and region. This resulted in the data being filtered to only individuals 18 or older. Two Cox models were developed for each cause of occurrence (models are provided). R packages were survival and cmprsk. The results of this study were that the Cox model showed that if the cause of the event was not considered, the risk based on gender did not vary. The competing risks model showed differently.

[@asghar2024] The goal of the paper is to examine the survival and risk of death associated with Covid-19 in the hospitals of a Pakistani city. Specifically, the paper looks at asymptomatic vs symptomatic individuals and focuses on the effects of gender, age, and if the person was symptomatic to assist in determining an individual's survival if they had the virus. This paper and research is important as Covid-19 demonstrated that a virus can and will still spread massively throughout the world. Determining the chances of survival with the virus would have been beneficial in the height of the pandemic, but is also beneficial now as we are not going to stop living with the virus anytime soon. The data the paper uses is from the Directorate General Health Services in KP, Pakistan and contains the hospital attendance records from the first wave of the pandemic, specifically March 2020 to June 2020. Individuals age, gender, date of illness, dates of admission and discharge/deceased, and if the individual had symptoms or not when diagnosed with Covid-19. The study focuses on individuals who are younger than 100 and did not have missing values. They first examined the Kaplan-Meier curve for the survival probability from their data. This allowed them to determine that the conventional Cox proportional hazards model may not be best since patients would recover and continue to survive, but the Cox model assumes survival probability will eventually go to zero. As such, they also determined a second model using logisitic regression and the Weibull regression model. They used a mixture cure model and a non-mixture model for the Weibull regression model. Using the Cox proportional hazards model, with three covariates (age, gender, being asymptomatic), age and being asymptomatic were considered to be significantly associated with survival of Covid-19 patients. Patients who are older and were symptomatic had a higher risk of death. However, the model showed that gender was insignificant. For the Weibull mixture cure (which allows for a patients to be cured), older and symptomatic patients were again at an increased risk of death if the person was uncured. However, those patients did have a longer survival rate than the uncured asymptomatic patients. For the Weibull non-mixture cure model, age and being asymptomatic indicated an increased risk of death in the long-term survival, with the symptomatic indicating an increased risk in the short-term survival. Gender for both Weibull models was not considered significant.

[@jiang2024] The goal of the paper is to critique another paper on chest pain incidence and mortality in the US and provide the limits of the model used in that paper. Heart disease is the leading cause of death for the US according to the CDC. As such, research into the topic, especially attempting to predict the event of death should be thorough and well thought out. This paper address concerns for a research topic done. The study the paper focuses on found that there is a significant relationship between Triglyceride-Glucose and mortality. This was regardless of chest pain occurring within the patients. However, there were several limitations using the Cox proportional hazards model as the study did. The first limitation brought up is that the Cox model assumes constant hazard ratios over time, but that the effect of Triglyceride-Glucose could vary over time which would violate the assumption of proportional hazards ratios. That violation could lead to biased estimates and inaccurate conclusions. The Cox model also struggles with time-dependent variables. The study took into account body weight and blood glucose levels, both of which can (and will) change throughout time which could impact the true effect of Triglyceride-Glucose. The paper provides several other weaknesses of the original study. The paper ens by suggesting a better alternative approach would have been to use time-varying covariate models. This paper was helpful as it provided data driven examples of why Cox proportional hazards model may not be the best model. This will be helpful in guiding the research into datasets to model.