### 2025-02-17 articleSummaries.qmd
Brad edits 2/16

---
title: "Literature Review - Reed"
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
bibliography: references.bib
---

[@ziadi2023]This article focused on the length of productive life (LPL in the article) for female Floridian goats. The article discusses the failures of linear models and random regression linear models when it comes to modeling the LPL. As such, Cox proportional hazards model was chosen because it can handle censored records, time-dependent variables, and skewed distribution of characteristics. The does were classified as censored if they were living and uncensored if they were deceased. A preliminary analysis using likelihood ratio tests on a Weibull hazard survival model. This showed that the effects selected all had a significant effect on LPL. The Cox model is then provided for this specific problem. The results were that the age of the first kidding did significant influence LPL. The article provides tables of their results.

[@zapletal2021] The goal of this paper was to examine how explanatory variables influence on an event and if the estimates of the effects of covariates will differ based on the approach that was selected. This paper in particular focused on the Cox proportional hazard and the competing risks approach. The Cox proportional hazards model provides an unspecified baseline hazard function which is beneficial when dealing with a single outcome of interest or one that does not have a specific cause (equations are provided in article). The article describes which competing risks model is the one that will be used in the modeling section. The predictors used by the author were gender, age when the contract began, and region where the person lived. The particular data used in this article was from a Czech Republic insurance company who provide policies for certain illness events until the event occurs. The insured will receive money if the event occurs, but their contract is then complete. Cox models have an assumption of proportional hazards which means the ratios of the hazards need to be independent of time to use the model. The test used was the zph test using Schoenfield residuals (paper is linked in article) and the hypothesis was rejected for age, but not for gender and region. This resulted in the data being filtered to only individuals 18 or older. Two Cox models were developed for each cause of occurrence (models are provided). R packages were survival and cmprsk. The results of this study were that the Cox model showed that if the cause of the event was not considered, the risk based on gender did not vary. The competing risks model showed differently.

[@asghar2024] The goal of the paper is to examine the survival and risk of death associated with Covid-19 in the hospitals of a Pakistani city. Specifically, the paper looks at asymptomatic vs symptomatic individuals and focuses on the effects of gender, age, and if the person was symptomatic to assist in determining an individual's survival if they had the virus. This paper and research is important as Covid-19 demonstrated that a virus can and will still spread massively throughout the world. Determining the chances of survival with the virus would have been beneficial in the height of the pandemic, but is also beneficial now as we are not going to stop living with the virus anytime soon. The data the paper uses is from the Directorate General Health Services in KP, Pakistan and contains the hospital attendance records from the first wave of the pandemic, specifically March 2020 to June 2020. Individuals age, gender, date of illness, dates of admission and discharge/deceased, and if the individual had symptoms or not when diagnosed with Covid-19. The study focuses on individuals who are younger than 100 and did not have missing values. They first examined the Kaplan-Meier curve for the survival probability from their data. This allowed them to determine that the conventional Cox proportional hazards model may not be best since patients would recover and continue to survive, but the Cox model assumes survival probability will eventually go to zero. As such, they also determined a second model using logisitic regression and the Weibull regression model. They used a mixture cure model and a non-mixture model for the Weibull regression model. Using the Cox proportional hazards model, with three covariates (age, gender, being asymptomatic), age and being asymptomatic were considered to be significantly associated with survival of Covid-19 patients. Patients who are older and were symptomatic had a higher risk of death. However, the model showed that gender was insignificant. For the Weibull mixture cure (which allows for a patients to be cured), older and symptomatic patients were again at an increased risk of death if the person was uncured. However, those patients did have a longer survival rate than the uncured asymptomatic patients. For the Weibull non-mixture cure model, age and being asymptomatic indicated an increased risk of death in the long-term survival, with the symptomatic indicating an increased risk in the short-term survival. Gender for both Weibull models was not considered significant.

[@jiang2024] The goal of the paper is to critique another paper on chest pain incidence and mortality in the US and provide the limits of the model used in that paper. Heart disease is the leading cause of death for the US according to the CDC. As such, research into the topic, especially attempting to predict the event of death should be thorough and well thought out. This paper address concerns for a research topic done. The study the paper focuses on found that there is a significant relationship between Triglyceride-Glucose and mortality. This was regardless of chest pain occurring within the patients. However, there were several limitations using the Cox proportional hazards model as the study did. The first limitation brought up is that the Cox model assumes constant hazard ratios over time, but that the effect of Triglyceride-Glucose could vary over time which would violate the assumption of proportional hazards ratios. That violation could lead to biased estimates and inaccurate conclusions. The Cox model also struggles with time-dependent variables. The study took into account body weight and blood glucose levels, both of which can (and will) change throughout time which could impact the true effect of Triglyceride-Glucose. The paper provides several other weaknesses of the original study. The paper ens by suggesting a better alternative approach would have been to use time-varying covariate models. This paper was helpful as it provided data driven examples of why Cox proportional hazards model may not be the best model. This will be helpful in guiding the research into datasets to model.

[@wang2025] The goal of the paper is to try and find a connection between the salt intake of an individual and if they will develop anxiety or depression later in life. Depression and anxiety are extremely prevalent in our society today and we are prone to eating salt with every meal, especially in the United States. Determining an association between salt and mental health will allow professionals an area to ask their patients to assist them with their mental health. In addition, salt is known to be harmful in large quantities to our health overall so this would be another area to encourage a decrease in salt intake. The group took almost half a million participants who, at the start of the study, had never been diagnosed with depression or anxiety and had no missing data. The participants were asked if they added salt to their foods and to note how often they do and major changes to their diet, if any. They first found the mean with SD and used the chi-square tests for categorical variables and general linear models for continuous variables to check the proportions between their participants and their salt intake level. They then used the Cox proportional hazards for the association between an individual's salt intake and the risk of developing depression or anxiety. They tested the assumption using the Kaplan-Meier method and Schoenfeld residuals and found no violation. They looked at the analyses with hazard ratios and did multiplicative interaction assessments in the Cox models. They also conducted several sensitivity tests to make adjustments. The results showed that hazard of depression did increase with the increased intake of salt. A similar result was found for anxiety and increased salt. The limitations of this particular study included the self-reporting of a person's salt intake, their sociodemographic factors, and their lifestyle factors. Primarily, the limitations for this paper were based on possible bias in their data.

[@zhang2025] The goal of the paper is to determine an association with postpartum depression and having two or more chronic diseases (called multimorbidity). This is important because many women suffer from postpartum depression and this can have lasting affects on their lives. In addition, understanding the long term implications of postpartum depression could help in preparing women for what might happen with childbirth and the mental impact it can have. The data was provided by self reporting. Women noted if they had a depressive episode in the months after giving birth or if they were told they had postpartum depression. They selected 36 chronic diseases to focus on and used the prevalence of multimorbidity, the incidence of multimorbidity, the number of new-onset chronic diseases, the cumulative rate of chronic diseases, and the multimorbidity patterns as the specific outcomes they were measuring. They provided the list of covariates within the paper. They first used binomial logistic regression models and quasi-Poisson mixed effects models to assist in gaining the initial findings of the study and create a baseline. They used Cox proportional hazard models to evaluate the associations of postpartum depression with having developed one or more of their chronic diseases. They checked the proportional hazard assumption using Schoenfeld's residual methods and no violations were found. They then used multinomial logistic regression models to study the association between postpartum depression and the five multimorbidity patterns they were able to determine. Patients who had a history of postpartum depression were shown to have a higher rate prevalence and incidence of one or more of the chronic diseases. They also discovered that those who had multimorbidity as their baseline prior to the study had a higher rate of prevalence and incidence as well. The limitations of this study are similar to those in the salt and anxiety study as their data came from the same place in both studies. As such, the data was open to bias and issues with self-reporting from the patients. The data also limited the representation of women due to various factors.

[@Ming-Chiang2014]

https://citeseerx.ist.psu.edu/document?repid=rep1&type=pdf&doi=8e45bb12b4bf296f002977caf11e969d5cc90a02

This article uses Cox Proportional Hazards model to predict bankruptcies in companies in the Taiwanese stock market. The model used uses 12 variables from companies found between 2003 and 2009 to classify companies into the categories of distressed vs non-distressed to indicate if they are at risk of bankruptcy. Each variable had reasoning and/or sources to explain the reasoning behind using them in the model along with statistical justification. The resulting model had an accuracy of 88%.

[@Bustan2018]

https://iopscience.iop.org/article/10.1088/1742-6596/1028/1/012230/pdf

This article describes a Cox Proportional Hazards Model to predict the probability of survival in breast cancer patients. The model uses six variables to make predictions and provides an explanation of the risk factors that are the most heavily weighted when deciding a patient's outlook. The article includes a table showing the survival probabilities of patients based on the values of their corresponding covariates. The article concludes the most important factor in surviving breast cancer is to catch it in as early of a stage as possible and recommends more outreach programs to encourage individuals to get themselves examined by a doctor regularly.

[@Kuitunen2021]

This article explores the proportionality assumption of cox proportional hazard models on arthroplasty research datasets. It accounts for over 1100 studies. The article concludes that the proportionality assumption cannot be ignored, as without it the results of the models created are unreliable. It is critical for the proportional hazards assumption to be tested properly, or inaccurate predictions will be made,

[@Smith2003]

In this article, Cox Proportional Hazard modeling was used for modeling time until event while accounting for other variables. The author explains how Cox PH models work mathematically and the strenths of Cox PH models. Cox PH models are able to accomodate incomplete or "censored" data which is prevalent in survival data. In many clinical trials, a patient death may not be observed because the trial period will end not allowing enough time for the event to occur. Health datasets were analyzed with hospitalization and demographic data to predict hospitalization time after inhaling oil well fire smoke. The model creates a probability for each individual's chances of being hospitalized.

[@Abeysekera2009]

This article summarizes the use of Shoenfeld's global test for testing assumptions of Cox PH models. It mentions that other methods for checking assumptions have an aspect of subjectivity to them, but this method shines because it does not. The proportional hazards assumption is the most important aspect of a dataset when trying to analyze data with a Cox PH model, and the Schoenfeld's global test is an effective method to test the assumption. The article includes visualizations and ways to interpret the visuals. The example provided in the paper shows the conclusion of which covariates were shown to be in violation of the assumption in a better way than typical methods. The article recommends Shoenfeld's global test as the ideal method for testing assumptions.

[@Seung2023]

https://www.elifecycle.org/archive/view_article?pid=lc-3-0-8

This article compares Kaplan-Meier survival analysis with Cox Proportional Hazards models. It summarizes how effective and necessary the models are in the health industry, and how important it is to pass the proportional hazards assumption to ensure an accurate model. It demonstrates the reason Cox models are so widely used over Kaplan-Meier because they're able to accommodate multiple covariates into predictions. The paper also explains how to handle datasets that break the assumptions of Cox PH models. It also goes over the time dependent regression models that can be used when the assumptions are broken. The author concludes that each model mentioned has its merits, but Cox PH models are the most popular models because of their flexibility and ability to include multiple covariates.