### 2025-03-24 CoxPractice.qmd
Version error fixed.

---
title: "Cox Proportional Hazards Practice"
format: html
editor: visual
self-contained: true
execute: 
  warning: false
  message: false
---

#### General Examination of Data

```{r}
#libraries
library(tidyverse)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
```

Dataset is from the Mayo Clinic Primary Biliary Cirrhosis data on Kaggle: <https://www.kaggle.com/datasets/jixing475/mayo-clinic-primary-biliary-cirrhosis-data>

```{r}
pbc <- read.csv('C:/Users/jayre/Desktop/Capstone/pbc.csv')

head(pbc, n = 5)
```

Columns:

-   id: case number
-   time: number of days between registration and the earlier of death, transplantion, or study analysis time in July, 1986
-   status: 0=alive, 1=liver transplant, 2=dead
-   trt: 1= D-penicillamine, 2=placebo
-   age: in days
-   sex: 0=male, 1=female
-   ascites: presence of ascites (0=no 1=yes)
-   hepato: presence of hepatomegaly (0=no 1=yes)
-   spiders: presence of spiders (0=no 1=yes)
-   edema: presence of edema (0=no edema and no diuretic therapy for edema; .5 = edema present without diuretics, or edema resolved by diuretics; 1 = edema despite diuretic therapy)
-   bili: serum bilirubin in mg/dl
-   chol: serum cholesterol in mg/dl
-   albumin: in gm/dl
-   copper: urine copper in ug/day
-   alk.phos: alkaline phosphatase in U/liter
-   ast: SGOT in U/ml
-   trig: triglicerides in mg/dl
-   platelet: per cubic ml / 1000
-   protime: prothrombin time in seconds
-   state: histologic stage of disease

```{r}
#sex needs to be 0,1, need no na's in the treatment column, and don't want the transplant category in status
pbc <- pbc %>% 
        mutate(sex = if_else(sex == "f", 1, 0)) %>%
        na.omit(trt) %>%
        filter(status != 1)

#switch status to be 1 (alive) or 2 (deceased)
pbc <- pbc %>% mutate(status = if_else(status == 0, 1, 2))

head(pbc, n = 5)

```

#### Kaplan-Meier Survival Curves and Log-Rate Tests

```{r}
#Generate the Kaplan-Meier survival curve and plot it with ggplot
#pull out time and status
t <- pbc[, "time"]
s <- pbc[, "status"]

km_fit <- survfit(Surv(t, s) ~ 1)

autoplot(km_fit) + theme_bw()
```

```{r}
#generate the probability of survival at certain times between registration and death 
summary(km_fit, times = c(1:5, 50, 100, 150*(1:10), 300*(5:15)))
```

```{r}
#check km with sex
sex <- pbc[, "sex"]
km_sex <- survfit(Surv(t, s) ~ sex)

autoplot(km_sex) + theme_bw()
```

Can use the log-rank test from the survival test to compare the curves between the two groups.

Hypothesis test:

-   $H_0$: in terms of survivability, there is no difference between the two groups
-   $H_1$: there is a survival differential b/w the two groups

Use $\alpha = 0.05$ as the comparison for the p-value

```{r}
#log rank test to see the impact based on sex
survdiff(Surv(t, s) ~ sex)
```

p-value = $0.03 < \alpha$ so we reject the hull hypothesis as there is enough evidence of a survival difference b/w males and females.

```{r}
pbc %>% summarise(mean = mean(age))
#mean age = 50 so check for <= 50 (0) and > 50 (1) 
pbc <- pbc %>% mutate(age2 = if_else(age > 50, 1, 0))

age <- pbc[, "age2"]

survdiff(Surv(t, s) ~ age)
```

p-value $< 0.005 < \alpha$ so we reject the null hypothesis as there is evidence of a difference in survival between those older than 50 and those younger than 50.

```{r}
#plot curve
km_age <- survfit(Surv(t, s) ~ age)
autoplot(km_age) + theme_bw()
```

```{r}
#check with treatment
trt <- pbc[, "trt"]
survdiff(Surv(t, s) ~ trt)
```

$p = 0.6 > \alpha$ so we fail to reject the null hypothesis as there is insufficient evidence of a difference in survivability b/w treatment groups.

#### Univariate Cox Models

(mostly for practice in understanding the models)

```{r}
#sex
coxph(Surv(time, status) ~ sex, data = pbc) %>%
  summary()
```

log-rank matches the test from above with $p = 0.03$

```{r}
#age
coxph(Surv(time, status) ~ age, data = pbc) %>%
  summary()
```

```{r}
#trt
coxph(Surv(time, status) ~ trt, data = pbc) %>%
  summary()
```

#### Multivariate Models

```{r}
m <- coxph(Surv(time, status) ~ age2 + sex + trt, data = pbc)

m %>% summary()
```

For all three tests (likelihood ratio, Wald, and logrank) $p < .005$ which is $< \alpha = 0.05$ so the overall model is significant and we reject the omnibus hypothesis (all $\beta$s are 0).

In the multivariate world, we can see that trt is not significant as the p-value is $> \alpha = 0.05$. However, age and trt are still significant as their p-values are $< \alpha$.

The HRs are:

-   age: 2.0781
-   sex: 0.6163
-   trt: 0.9650

The HR for age and the lower p-value indicate a strong relationship b/w the patient's age and increased risk of death. Specific wording: holding the other covariates constant, an increased age is associated with a poor survivall.

#### Testing Assumptions

```{r}
#proportional hazards
cox.zph(m)
ggcoxzph(cox.zph(m))
```

All p-values are $> \alpha$ so we do not have any coefficients with that are still time dependent.

```{r}
#non-linearity
m_resid <- residuals(m, type = "martingale")

ggplot(data = pbc, mapping = aes(x = age2, y = m_resid)) + 
  geom_point() + 
  geom_smooth() + 
  theme_bw()
```

```{r}
#use the continuous age for comparison
ggplot(data = pbc, mapping = aes(x = age, y = m_resid)) + 
  geom_point() + 
  geom_smooth() + 
  theme_bw()
```

The continuous version of age would be in violation, but the categorical one is not.