### 2025-04-13 slides.qmd
Slides Update 4/13

---
title: "Predicting Survival Probability of NASA Aircraft Engines"
subtitle: "Using Cox Proportional Hazards Model"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: last-modified
date-format: long
format:
  revealjs:
    theme: moon
    smaller: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
  echo: false
editor: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE, warning=FALSE, include=FALSE}
#Load packages
library(tidyverse)
library(gridExtra)
library(grid)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)
library(broom)
```

```{r, echo = FALSE}
#generation of the model for sample plots
example_surv_object <- Surv(time = lung$time, event = lung$status)

example_coxmodel <- coxph(example_surv_object ~ meal.cal, data = lung)
```

## Cox Proportional Hazards (CPH) Model

What is it?

-   A statistical regression method specializing in modeling
    time-to-event predictions with survival data [@Abeysekera2009]
    -   Survival data has a value for time and an indicator column for
        an event
-   Is a method that can deal with censored data
    -   Censored data is when the information about an individual in a
        study is only known for a certain period of time [@klein2005]
-   Primarily used in the health field but has applications in
    predicting bank failure, the survival probability of machines, and
    insurance likelihood payouts

## Limitations

-   The model assumes that as time goes on, the survival probability
    will approach zero with no survivors [@asghar2024]

-   The proportional hazards assumption can limit the ability to
    correctly predict the effect of a variable [@jiang2024]

-   The covariate selection can become biased and may not accurately
    represent the true data [@wang2025],[@zhang2025]

-   The model cannot provide a specific value for when the event will
    happen, only the probability of when the event might happen

## Mathematical Formulas

-   Survival Function:

    $S(x) = \int_{x}^{\infty} f(x)dx$

-   Hazard Function:

    $h(x) = \lim\limits_{\Delta x \rightarrow 0} \frac{P[x \leq X < x + \Delta x|X \geq x]}{\Delta x}$

-   CPH Model Hazard Function:

    $h(t|\mathbf Z) = h_0(t)\text{exp}(\sum\limits_{k=1}^{p} \beta_kZ_k)$

-   Proportional Hazards Ratio:

    $\frac{h(t|\mathbf Z)}{h(t|\mathbf Z*)} = \text{exp}[\sum\limits_{k=1}^{p} \beta_k(Z_k - Z_k^*)]$

-   Cumulative Hazard Function:

    $H(x) = \int_0^x h(u) du$

## Assumptions

There are four assumptions for CPH:

::: incremental
-   Independence assumption

    -   Assumes that the survival times of observed subjects are
        independent of each other [@nahhas2025]

-   Non-informative Censoring Assumption

    -   Assumes that censoring is non-informative [@nahhas2025]
:::

## Assumptions, Cont.

-   Linearity Assumption

    ::: incremental
    -   Assumes the relationship between covariates and the outcome is a
        linear relationship

    -   This is tested using the Martingale residuals for each covariate
        with the equation Martingale Residuals = Observed Events -
        Expected Results [@nahhas2025]

        -   The residuals are examined visually

        -   If linear and appear to have a slop of zero, the assumption
            is not violated [@Amini2015]
    :::

```{r,echo=FALSE}
example_martingale <- resid(example_coxmodel, type = "martingale")

plot(example_martingale, 
     main = "Martingale Residuals: Example Covariate",
     ylab = "Martingale Residuals for Example Covariate",
     xlab = "Value Range for Example Covariate")
abline(h = 0, col = "red")

```

## Assumptions, Cont.

-   Proportional Hazards Assumption

    ::: incremental
    -   The ratio of hazard rates for any two subjects must be constant
        at all times [@Bustan2018]

    -   To test this assumption, the below hypothesis test is used:

        $$
        \begin{align*}
        H_0&:\; \text{the log hazard ratio for each covariate remains constant over time} \\
        H_1&: \; \text{the log hazard ratio for each covariate does not remain constant over time}
        \end{align*}
        $$

        -   For this test, the goal is to *not* reject $H_0$

    -   The scale Schoenfeld partial residuals are used to find the
        p-value for this test

        -   The Schoenfeld residuals are the difference between the
            value of the covariate and the expected value of the
            covariate at the time of failure [@klein2005]

        -   There will be multiple p-values generated: global and
            individual

            -   If the global p-value is $< 0.05$, then the model has at
                least one covariate that violates the proportional
                hazards assumption and the individual p-values must be
                examined

            -   If the global p-value is $> 0.05$, then all covariates
                meet the proportional hazard assumption [@nahhas2025]

    Example of Schoenfeld plot and results

    ```{r, echo = FALSE}
    example_phtest <- cox.zph(example_coxmodel)

    plot(example_phtest[1],
         ylab = "Covariate Schoenfeld Residual",
         xlab = "Time")

    multi <- coxph(example_surv_object ~ meal.cal + age, data = lung)
    ph_tibble <- cox.zph(multi) %>%
      pluck("table") %>%
      as.data.frame() %>%
      tibble::rownames_to_column("variable") %>%
      as_tibble()
    kable(ph_tibble, format = "markdown")

    ```
    :::

## Model Visualization - Kaplan Meier

## Model Visualization - Forest Plot

## Evaluation

## Prediction

## Data Structure, Exploration, and Visualization

## Creating CPH Model

## Checking Assumptions

## Adjusting the Model

## Model Results

## Conclusion

## References