### 2025-04-15 slides.qmd
Update 4/15 pt2

---
title: "Predicting Survival Probability of NASA Aircraft Engines"
subtitle: "Using Cox Proportional Hazards Model"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: last-modified
date-format: long
format:
  revealjs:
    theme: moon
    smaller: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
  echo: false
editor: 
  markdown: 
    wrap: 72
---

```{r, warning=FALSE, include=FALSE}
#Load packages
library(tidyverse)
library(gridExtra)
library(grid)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)
library(broom)
```

## Cox Proportional Hazards (CPH) Model

What is it?

::: incremental
-   A statistical regression method specializing in modeling
    time-to-event predictions with survival data [@Abeysekera2009]

    -   Survival data has a value for time and an indicator column for
        an event

-   Is a method that can deal with censored data

    -   Censored data is when the information about an individual in a
        study is only known for a certain period of time [@klein2005]

-   Primarily used in the health field but has applications in
    predicting bank failure, the survival probability of machines, and
    insurance likelihood payouts
:::

## Limitations

::: incremental
-   The model assumes that as time goes on, the survival probability
    will approach zero with no survivors [@asghar2024]

-   The proportional hazards assumption can limit the ability to
    correctly predict the effect of a variable [@jiang2024]

-   The covariate selection can become biased and may not accurately
    represent the true data [@wang2025],[@zhang2025]

-   The model cannot provide a specific value for when the event will
    happen, only the probability of when the event might happen
:::

## Mathematical Formulas

-   Survival Function: $S(x) = \int_{x}^{\infty} f(x)dx$

-   Hazard Function:
    $h(x) = \lim\limits_{\Delta x \rightarrow 0} \frac{P[x \leq X < x + \Delta x|X \geq x]}{\Delta x}$

-   CPH Model Hazard Function:
    $h(t|\mathbf Z) = h_0(t)\text{exp}(\sum\limits_{k=1}^{p} \beta_kZ_k)$

-   Proportional Hazards Ratio:
    $\frac{h(t|\mathbf Z)}{h(t|\mathbf Z*)} = \text{exp}[\sum\limits_{k=1}^{p} \beta_k(Z_k - Z_k^*)]$

-   Cumulative Hazard Function: $H(x) = \int_0^x h(u) du$

-   Concordance Index: $C = \frac{c + \frac{t_x}{2}}{c + d + t_x}$

-   Survival Probability: $S(t) = e^{-H(t)}$, where $H(t)$ is the above
    cumulative hazard function

## Assumptions

There are four assumptions for CPH:

::: incremental
-   Independence assumption

    -   Assumes that the survival times of observed subjects are
        independent of each other [@nahhas2025]

-   Non-informative Censoring Assumption

    -   Assumes that censoring is non-informative [@nahhas2025]

-   Linearity Assumption

    -   Assumes the relationship between covariates and the outcome is a
        linear relationship [@nahhas2025]

-   Proportional Hazards Assumption

    -   Assumes the ratio of hazards rates for any two subjects are
        constant at all times [@Bustan2018]
:::

## Visualization

The two common model visualizations are:

::: incremental
-   Kaplan-Meier Survival Curve

    -   Demonstrates the time where the event being modeled is expected
        to occur [@klein2005]

-   Forest Plots

    -   Visualizes the effects of each covariate in the model on the
        hazard ratio
:::

## Evaluation and Survival Probability

::: incremental
-   Model accuracy evaluation is done using the concordance index

    ::: incremental
    -   The concordance index measures the amount of agreement between
        two variables

    -   A value of 1 means all the pairs are correctly ordered while a
        value of 0 means no pairs are correctly ordered
    :::

-   Survival probability can be predicted at a specific time $t$

    ::: incremental
    -   If the probability is $\geq 50$%, it is assumed the event has
        not occurred

    -   If the probability is $< 50$%, it is assumed the event has
        occurred
    :::
:::

## Data Structure

The data selected for this project comes from a study on propagation
modeling that NASA completed, specifically focusing on the engine two
testing and training datasets [@saxena2008].

-   Each engine in the NASA data has an unknown amount of wear,
    manufacturing variation, and sensor noise

-   There are three operation setting fields and twenty-one sensor
    measurement fields

-   A column indicating status was added to both the testing and
    training datasets with 0 indicating the machine has not failed and 1
    indicating the machine has failed

-   The data was then merged together to make one grouped dataset

```{r}
testurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/test_FD002.txt'

test <- read.delim(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0,
         id = id + 260)

trainurl <- 'https://raw.githubusercontent.com/JR-87/capstone4ds_reed/refs/heads/main/train_FD002.txt'

train <- read.delim(trainurl, header = TRUE, sep = '') %>%
  group_by(id) %>%
  mutate(status = if_else(row_number() == n(), 1, 0)) %>%
  ungroup()

data_all <- rbind(train, test)

train_grouped <- train %>% 
  group_by(id) %>%
  slice_max(time)

data_grouped <- data_all %>%
  group_by(id) %>%
  slice_max(time)

temp <- data_grouped %>% select(id, time, status, os1, os2, os3, sm1, sm2, sm3)

kable(head(temp, n = 5), format = "markdown", caption = 'NASA Aircraft Engine Data')
```

## Exploration

-   519 engines in the combined data

    -   260 engines in training data

    -   259 engines in testing data

```{r}
summarymetrics <- data.frame(c("Minimum", "Median", "Mean", "Standard Deviation", "Maximum"),
                             c(min(train_grouped$time), 
                               median(train_grouped$time),
                               mean(train_grouped$time),
                               sd(train_grouped$time),
                               max(train_grouped$time)))

colnames(summarymetrics) <- c("Metric", "Value")
summarymetrics$Value <- round(summarymetrics$Value, digits = 2)
kable(summarymetrics, format = "markdown", align = "c", caption = "Summary Metrics")
```

## Visualization

::: panel-tabset
#### Histogram/Boxplot

```{r}
#| fig-align: center
Histogram <- ggplot(data = train_grouped,
       aes(x = time)) +
  geom_histogram(binwidth = 5,
                 fill = "grey")+
  labs(x = "Survival Time (Iterations)",
       y = "Frequency")+
  theme_bw()

box <- ggplot(data = train_grouped,
                  aes(y = time))+
  geom_boxplot(fill = "gray",
               color = "black")+
  ylab("Time (Iterations)")+
  theme_bw()

grid.arrange(Histogram,box,ncol = 2, top = "Distribution of Engine Longevity in Training Data")
```

#### Survival Curve

```{r}
#| fig-align: center
t <- as.matrix(train_grouped[, "time"])
s <- as.matrix(train_grouped[, "status"])

km_fit <- survfit(Surv(t, s) ~ 1)

autoplot(km_fit) + theme_bw()
```
:::

## Creating CPH Model

The table provides the model number, the covariates used, the AIC and
BIC from the stepwise regression if applicable, and the concordance
index.

```{r}
surv_object <- Surv(data_grouped$time, data_grouped$status)

all_cov <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

aic <- MASS::stepAIC(all_cov, trace = 0)
n <- nrow(data_grouped)
bic <- step(all_cov, direction = "both", k = log(n), trace = 0)

cox1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

cox2 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm5 + sm8 + sm9 + sm15 + sm16 + sm17 + sm18 + sm20)

cox3 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm8 + sm9 + sm15 + sm18)

cox4 <- coxph(data = data_grouped, surv_object ~ os3 + sm13 + sm14 + sm19)

stepwise_results <- data.frame(c("model 1", "model 2", "model 3", "model 4"), 
                               c("All covariates", "os3, sm3 - sm5, sm8 - sm9, sm15 - sm18, sm20", "os3, sm3 - sm4, sm8-sm9, sm15, sm18", "os3, sm13 - sm14, sm19"),
                               c("2449.92", "2431.88", "N/A", "N/A"),
                               c("2547.71", "N/A", "2461.31", "N/A"),
                               c(cox1$concordance[6], cox2$concordance[6], cox3$concordance[6], cox4$concordance[6]))

colnames(stepwise_results) <- c("Model", "Covariates", "AIC", "BIC", "Concordance")

kable(stepwise_results, format = "markdown")
```

Model for the continuing analysis will have all covariates except sm16
and sm19.

## CPH Model Forest Plot

```{r}
#| fig-align: center
cox1.1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm17 + sm18 + sm20 + sm21)

ci <- cox1.1$concordance[6]

plot_model(cox1.1,
           dot.size = 2,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")
```

## Checking Assumptions

## Model Results

## Conclusion

## References